# Reproducible Research: Peer Assessment 1




## Loading and preprocessing the data
*Place the data file archive in a folder called ' data ' under the current directory*
```{r}
classes = c("integer", "Date", "integer")
activity.df <- read.csv(unz("data/repdata_data_activity.zip", filename = "activity.csv"), colClasses = classes)
activity.complete.df <- na.omit(activity.df)
```





## What is mean total number of steps taken per day?

```{r}
x <- with(activity.complete.df, tapply(steps, as.factor(date), sum))
hist(x, breaks = 10, col = "yellow")
```
Mean number of steps taken per day : `r mean(x)`  
   
Median of steps taken per day : `r median(x)`




## What is the average daily activity pattern?

```{r}
intv <- as.factor(activity.complete.df[, "interval"])
intv.mean <- tapply(activity.complete.df[, "steps"], intv, mean)
plot(x = unique(intv), y = intv.mean, type = "l", cex = 2)
```

Interval with the highest average number of steps : `r rownames(intv.mean)[which(intv.mean == max(intv.mean))]`

## Imputing missing values
Total number of missing values in the dataset = `r nrow(activity.df) - nrow(activity.complete.df)`

Replace missing value for a given interval on a certain date with the mean value of steps for that interval across all states
```{r}
for(i in 1 : nrow(activity.df)) {
    if(is.na(activity.df[i, "steps"]))  
        activity.df[i, "steps"] <- intv.mean[as.character(activity.df[i, "interval"])]
}
x <- with(activity.df, tapply(steps, as.factor(date), sum))
hist(x, breaks = 10, col = "yellow")
```
Mean number of steps taken per day : `r mean(x)`  
   
Median of steps taken per day : `r median(x)`



## Are there differences in activity patterns between weekdays and weekends?
```{r}

daytype <- function(date.in) {
  
  err.date.handler <- function(expr) stop(paste(expr, " not a valid 'date' "), call. = F)
    
  day <- tryCatch(weekdays(as.Date(date.in), abbreviate = T), error = function(e) err.date.handler(date.in))
  if(day == "Sat" | day == "Sun")
      return("weekend")
  else
      return("weekday")
    
}

activity.df <- cbind(activity.df, "datetype" = sapply(activity.df[, "date"], daytype))
l <- split(activity.df, activity.df[, "datetype"])

# Calculate weekday average of steps
weekday.mean <- tapply(l[["weekday"]][, "steps"], as.factor(l[["weekday"]][, "interval"]), mean)
# Calculate weekend average of steps
weekend.mean <- tapply(l[["weekend"]][, "steps"], as.factor(l[["weekend"]][, "interval"]), mean)  

u <- data.frame("interval" = rownames(weekend.mean), "mean.steps" = as.vector(weekend.mean), "datetype" = "weekend")
v <- data.frame("interval" = rownames(weekday.mean), "mean.steps" = as.vector(weekday.mean), "datetype" = "weekday")
alldays.df <- rbind(u, v)


library(lattice)

xyplot(mean.steps ~ interval | datetype, data = alldays.df, layout = c(1, 2), type = "l", xlab = "Interval", ylab = "Steps")
```
